---
layout: page
title: blah
output:
  github_document:
    pandoc_args:  --webtex
  #  md_document:
#    variant: markdown_github

  preserve_yaml: true
bibliography: [../references/attacks.bib]
csl: [../references/apa-6th-edition.csl]
---

```{r , echo = FALSE, eval = TRUE}
knitr::opts_knit$set(base.dir = "/home/rafal/UG/UGprojects/redditAttacks",
base.url = "https://rfl-urbaniak.github.io/redditAttacks/")
knitr::opts_chunk$set(fig.path = "images/")
```




```{r setup, include=FALSE, echo=FALSE}
require("knitr")
#opts_knit$set(root.dir = "../")
options(knitr.kable.NA = '')

library(tidyverse)
library(knitr)
library(kableExtra)
library(formatR)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 60), tidy = TRUE)
```







\noindent \textbf{Abstract.} 
We conduct a large scale data-driven analysis of the effects of online personal attacks on social media user activity. First, we perform a thorough overview of the literature on the influence of social media on user behavior, especially on the impact that negative and aggressive behaviors, such as harassment and cyberbullying, have on users’ engagement in online media platforms. The majority of previous research were  small-scale self-reported studies, which is their limitation. This motivates our data-driven study. We perform a large-scale analysis of messages from Reddit, a discussion website, for a period of two weeks, involving 182,528 posts or comments to posts by 148,317 users. To efficiently collect and analyze the data we apply a high-precision personal attack detection technology. We analyze the obtained data from three perspectives: (i) classical statistical methods, (ii) Bayesian estimation, and (iii) model-theoretic analysis. The three perspectives agree: personal attacks decrease the victims’ activity.
The results can be interpreted as an important signal to social media platforms and policy makers that leaving personal attacks unmoderated is quite likely to disengage the users and in effect depopulate the platform. On the other hand, application of cyberviolence detection technology in combination with various mitigation techniques could improve and strengthen the user community. As more of our lives is taking place online, keeping the virtual space inclusive for all users becomes an important problem which online media platforms need to face.



\textbf{Remark.} In what follows, we sometimes display key pieces of code and explain what it does. Some not too fascinating pieces of code are supressed, but the reader can look them up in the associated .Rmd file and compile their own version.





#Technology applied for personal attack detection



For the need of this research we define personal attack as any kind of abusive remark made in relation to a person (ad hominem) rather than to the content of the argument expressed by that person in a discussion. The definition of ‘personal attack’ subsumes the use of specific terms which compare other people to animals or objects or making nasty insinuations without providing evidence. Three examples of typical personal attacks are as follows.


- *You are legit mentally retarded homie.*
- *Eat a bag of dicks, fuckstick.*
- *Fuck off with your sensitivity you douche.* 


The detection of personal attacks was performed using  Samurai, a proprietary technology of Samurai Labs.^[[https://www.samurailabs.ai/](https://www.samurailabs.ai/), described in [@ptaszynski2018cyberbullying;@wroczynski2019system].]


